An efficient acyclic contact planner for multiped robots

### 2016-12-12 body/intro.tex
rewrote intro

Reaching reasonable computation times is critical for a contact planner to be of any practical use. Our work thus aims at breaking the complexity of the contact planning problem.  To do so we deal sequentially with the two main issues associated to our problem: the null measure of the contact manifold, and the non-determinism of the contact selection problem. First we introduce a low dimensional space, called the \textit{contact-feasible} space, that can be sampled and mapped efficiently to the contact manifold. Then, given a path computed in the \textit{contact-feasible} space, we propose a deterministic algorithm to generate a contact sequence along the path.
This decoupling between the planning of the path and the generation of contacts presents drawbacks and advantages discussed in previous related literature.
%~ whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
%~ We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
%~ of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
%~ We then use a deterministic algorithm
%~ This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
%~ This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

@@

The issue of planning acyclic contacts was first completely described by Bretl et al. in their seminal paper  \citep{Bretl:2006:MPM:1124573.1124585}. The issue requires the simultaneous handling of two problems, $\mathcal{P}_1$: planning a relevant guide path for the root of the robot in $SE(3)$; and $\mathcal{P}_2$: planning a discrete sequence of acyclic equilibrium configurations along the path. A third nontrivial problem, $\mathcal{P}_3$, not addressed in this work, then consists in interpolating a complete motion between two postures of the contact sequence.  A key issue is to avoid combinatorial explosion when considering at the same time the possible contacts and the potential paths. This seminal paper proposes a first effective algorithm, able to handle simple situations (such as climbing scenarios), but not applicable to arbitrary environments. Following it, seve\-ral papers have applied this approach in specific situations, limiting the combinatorial by imposing a fixed set of possible contacts \citep{Hauser06usingmotion, stilman2010}.

@@

Our solution belongs the class of decoupled approaches, because we believe that the separation between the generation of the root guide path and the contact sequence is the most promising direction \citep{DBLP:conf/iser/EscandeKMG08} to break the complexity of multi-contact planning. Compared to previous approaches, our solution has two main novelties:

\noindent \textbf{Regarding $\mathcal{P}_1$}, we define and use the \textit{reachability condition} to plan the guide path. The \textit{reachability condition} defines a low dimensional space that can be mapped to the contact manifoldcomputationally-efficient approximation of the \glslink{equilibrium feasible}{\textit{equilibrium feasibility}}. This step is fast because the planning happens
in a low-dimensional space, and does not require explicit contact computations.

\noindent \textbf{Regarding $\mathcal{P}_2$},  we address \Pb by extending the path into a sequence of whole-body configurations in static equilibrium. This second step requires the explicit computation of contact configurations. It is guided by dedicated heuristics which fastly synthesize relevant configurations.

This sequential approach is the key to the efficiency of our method, though it
can result in failures because our hypothesis is not always true. However, we demonstrate empirically the validity of the approach: the high success rate, combined with the low computation times, allow us to plan (and re-plan upon failure) multi-contact sequences at \gls{interactive} rates.

@@

>>>>

Obtaining reasonable computation times is critical for a contact planner to be of any practical use. Our work thus aims at breaking the complexity of the acyclic contact planning problem.  To do so we deal sequentially with the two main issues associated to our problem: the null measure of the contact manifold, and the non-determinism of the contact selection problem. First we introduce a low dimensional space, called the \textit{contact-feasible} space, that can be sampled and mapped efficiently to the contact manifold. Then, given a path computed in the \textit{contact-feasible} space, we propose a deterministic algorithm to generate a contact sequence along the path.
This decoupling between the planning of the path and the generation of contacts presents drawbacks and advantages discussed in previous related literature.
%~ whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
%~ We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
%~ of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
%~ We then use a deterministic algorithm
%~ This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
%~ This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

@@

The issue of planning acyclic contacts was first completely described by Bretl et al. in their seminal paper  \citep{Bretl:2006:MPM:1124573.1124585}. The issue requires the simultaneous handling of two problems, $\mathcal{P}_1$: planning a relevant guide path for the root of the robot in $SE(3)$; and $\mathcal{P}_2$: planning a discrete sequence of acyclic equilibrium configurations along the path. A third nontrivial problem, $\mathcal{P}_3$, 
%~ not addressed in this work, 
then consists in interpolating a complete motion between two postures of the contact sequence.  A key issue is to avoid combinatorial explosion when considering at the same time the possible contacts and the potential paths. This seminal paper proposes a first effective algorithm, able to handle simple situations (such as climbing scenarios), but not applicable to arbitrary environments. Following it, seve\-ral papers have applied this approach in specific situations, limiting the combinatorial by imposing a fixed set of possible contacts \citep{Hauser06usingmotion, stilman2010}.

@@

Our solution belongs the class of decoupled approaches, because we believe that the separation between the generation of the root guide path and the contact sequence is the most promising direction \citep{DBLP:conf/iser/EscandeKMG08} to break the complexity of acyclic contact planning. Compared to previous approaches, our solution has two main novelties:

\noindent \textbf{Regarding $\mathcal{P}_1$}, we propose a fast guide path planning algorithm. The key to its efficiency is that it does not sample
directly the contact manifold, but an approximation of the \textit{contact feasible} space, for which there exists a mapping to the contact manifold.
The space is sampled thanks to the \textit{reachability condition} we introduce, a low-dimensional geometric condition for \textit{contact feasibility}.

\noindent \textbf{Regarding $\mathcal{P}_2$},  we propose a fast method to extend a \textit{contact feasible} path into a sequence of whole-body configurations in static equilibrium. This second step requires the explicit computation of contact configurations. It is guided by dedicated heuristics which fastly synthesize relevant configurations.

This sequential approach is the key to the efficiency of our method, though it
can result in failures:  we show that the \textit{reachability condition} only approximates the \textit{contact feasible} space. However, we demonstrate empirically the validity of the approach: the high success rate, combined with the low computation times, allow us to plan (and re-plan upon failure) multi-contact sequences at \gls{interactive} rates.

@@

Finally, to demonstrate the validity of our approach regarding the complete chain, we show that the contact plans we produce can be successfully executed, either in simulation or on the real HRP-2 robot.
